---
title: "Testing group differences"
output: html_document
editor_options: 
  chunk_output_type: console
---

A t-test can be done using paired data and unpaired data. The first example is an analysis of paired data. A key point is that we infer about the population with inferential statistics (based on a t-test) and describe our data with descriptive statistics (e.g. figures).

# A paired example

Here we are interested in the difference in CSA between slow and fast muscle fibers.

```{r load-data}
# Load variables
library(tidyverse)

hypertrophy <- read_csv("./data/hypertrophy.csv")
# read_csv is from the package readr


# Save the subsetted data set 
dat1 <- hypertrophy %>%
  select(SUB_ID, FAST_CSA_T1, SLOW_CSA_T1) 


# The t-test is done using the $ operator, selecting a specific column 
# in the data set. Save the t-test for later use
paired.t <- t.test(dat1$FAST_CSA_T1, dat1$SLOW_CSA_T1, 
       paired = TRUE)


# Combine code like so:

# The results can be written in the text as: The mean difference between 
# fibertypes were `r round(paired.t$estimate, 2)`, 95% CI: 
# [`r round(paired.t$conf.int[1], # 2)`, `r round(paired.t$conf.int[2], 2)`]. 



```

The results can be written in the text as: The mean difference between fibertypes were `r round(paired.t$estimate, 2)`, 95% CI: [`r round(paired.t$conf.int[1], 2)`, `r round(paired.t$conf.int[2], 2)`]. 

A descriptive plot can use every individual value or a combination. Below we see every paired value.

```{r plot-data}
dat1 %>%
  pivot_longer(names_to = "fibertype", 
               values_to = "csa", 
               cols = FAST_CSA_T1:SLOW_CSA_T1) %>%
  mutate(fibertype = gsub("_CSA_T1", "", fibertype)) %>%
  
 
  
  ggplot(aes(fibertype, csa, group = SUB_ID)) + 
  geom_line() +
  geom_point()


```

# An independent example

Independent t-tests are performed when the data have no relation, as with independent groups. 

```{r}

# Make a sub set of the data set

dat1 <- hypertrophy %>%
  select(SUB_ID, CLUSTER, T3T1__TESTOSTERONE_ngdl) %>%
  filter(!is.na(CLUSTER), 
         !is.na(T3T1__TESTOSTERONE_ngdl)) 
# The t test is unpaired
test <- t.test(T3T1__TESTOSTERONE_ngdl ~ CLUSTER, 
               data = dat1)


# Save data in code chunk for printing in text.

t <- round(test$statistic, 2)
df <- round(test$parameter, 1)
p <- round(test$p.value, 3)

estimate <- round(test$estimate[1] - test$estimate[2], 1)

cil <- round(test$conf.int[1], 1)
ciu <- round(test$conf.int[2], 1)

# To show these data in the text use the following code: 

# We conclude that HIGH does not differ from LOW
# in Testo change from T1 to T3 (mean
# difference, `r estimate`, 95% CI [`r cil`, `r
# ciu`]; t(`r df`) = `r t`, p = `r p`)


```

We conclude that HIGH does not differ from LOW in Testo change from T1 to T3 (mean difference, `r estimate`, 95% CI [`r cil`, `r ciu`]; t(`r df`) = `r t`, p = `r p`)














